\vspace{-0.25em}\thetitle

### 2009-07-31 paper.tex
...

The rest of this paper is organized as follows.
In \Section{background}, we provide some background and the present the observations which motivate this work.
Our model and hypothesis about flow behavior is presented in \Section{hypothesis}.
The experimental methodology for testing this hypothesis is presented in \Section{methodology} while the results follow in \Section{results}.
\Section{discussion} discusses the meaning and impact of these results and
\Section{related-work} gives context for this research in relation prior work.
We conclude with final remarks in \Section{conclusions}.

@@

\section{Discussion}\sec{discussion}

@@

While these potential applications are useful, the real significance of this work lies elsewhere.
Predicting flow behavior is merely a pleasant side effect of the real breakthrough:
a detailed statistical model for network-wide flow behaviors.
The fundamental concept underlying this work is that there are a relatively small collection of ``basic behaviors'' that generate the vast majority of the flow behaviors found in any particular trace.
These behaviors are associated with distributions of various flow features:
packet sizes, inter-packet intervals, \caps{IP} protocol, port numbers and packet counts.
These features are drawn from these distributions to create actual flows.
% We have treated the features as though they are sampled independently, but no doubt they are not really independent at all.

Given that the statistical model we have proposed for flow behavior is fairly simple---and finite mixture modeling is a well established field that has been studied since 1894~\cite{McLachlan00}.
Traditional approaches, however, have only applied to parametric probability distributions.
The distributions found in network traffic, however, are very distinctly non-parametric.
They are lumpy, unruly and generally quite ugly.
This is due largely to the fact that behaviors found in computer networks are not the by-products of smooth, natural phenomena, but rather of man-made artificial systems, following arbitrary, complex and capricious rules.

@@

It is our hope and belief that this model of flow behavior will provide future insights into other aspects of network traffic modeling:
in particular traffic classification and realistic workload generation.
The recovered model used in this paper is very close to providing a usable model for traffic generation already.
The largest missing piece is a model for realistic weight matrices.
If such applications based on this flow behavior model or other similar ones prove as fruitful as the application of flow behavior prediction, that will provide further evidence of the effectiveness and validity of our proposed model.

@@

\vfill
\pagebreak

\section{Related Work}\sec{related-work}

>>>>

The rest of this paper is organized as follows.
In \Section{background}, we provide some background and the present the observations which motivate this work.
Our model and hypothesis about flow behavior is presented in \Section{hypothesis}.
The experimental methodology for testing this hypothesis is presented in \Section{methodology} while the results follow in \Section{results}.
\Section{discussion} discusses the meaning and impact of these results and gives context for this research in relation prior work.
We conclude with final remarks in \Section{conclusions}.

@@

\section{Discussion \& Related Work}\sec{discussion}\sec{related-work}

@@

While these potential applications are useful, the real significance of this work lies elsewhere.
Predicting flow behavior is merely a pleasant side effect of the real breakthrough:
a detailed statistical model for network-wide flow behaviors.
The fundamental concept underlying this work is that there are a relatively small collection of ``basic behaviors'' that generate the vast majority of the flow behaviors found in any particular trace.
These behaviors are associated with distributions of various flow features:
packet sizes, inter-packet intervals, \caps{IP} protocol, port numbers and packet counts.
These features are drawn from these distributions to create actual flows.
We have treated the features as though they are sampled independently, but this is merely a simplification that allows us to estimate model parameters with computational techniques available today.
The fact that this estimation provides a model which can reasonably well predict the behaviors of flows demonstrates that the concept of modeling flows as a mixture of basic behaviors is viable, even if the details of how we represent flow features remains open to improvement.

Given that the statistical model we have proposed for flow behavior is fairly simple---and finite mixture modeling is a well established field that has been studied since 1894~\cite{McLachlan00}---one may very well ask why this basic linear structure of network traffic has not been discovered before.
Part of the reason is that traditional approaches to mixture modeling have been applicable to parametric probability distributions.
The distributions found in network traffic, however, are very distinctly non-parametric.
They are lumpy, unruly and generally quite ugly.
This is due largely to the fact that behaviors found in computer networks are not the by-products of smooth, natural phenomena, but rather of man-made artificial systems, following arbitrary, complex and capricious rules.

@@

It is our hope and belief that this model of flow behavior will provide future insights into other aspects of network traffic modeling:
in particular traffic classification and realistic workload generation.
The recovered model used in this paper is very close to providing a usable model for traffic generation already.
The largest missing piece is a model for generating realistic weight matrices.
With realistic basic behavior matrices extracted from traces and a means of generating realistic weight matrices, we have immediately a model for generating realistic traffic traces.
Earlier research has show that it is sufficient for simulation purposes to generate flows with realistic packet size distributions, inter-packet interval distributions, and packet counts by repeatedly sampling packets from these distributions~\cite{Karpinski07:realism,Karpinski07:cbr-failure}.
If such applications based on this flow behavior model or other similar ones prove as fruitful as the application of flow behavior prediction, that will provide further evidence of the effectiveness and validity of our proposed model.

@@
